### 2021-04-09 final-report.latex/main.tex
add theta term derivation, autocorrelation time and thermalization

\section{Monte Carlo Simulations}
\label{sec:mc}
We implement a Markov Chain Monte Carlo method following Schaich's thesis \cite{schaich2006}. This implementation utilizes a ``random walk,'' i.e. a set of random steps through phase space, to determine statistical values such as correlation functions across the lattice. By the definition of the Markov chain, the probability of adoption of each state, and therefore its inclusion in the Monte Carlo calculation, depends only on the current state and the proposed state. This probability is denoted as $P(\mu\rightarrow\nu)$ where $\mu$ and $\nu$ are the existing and proposed lattice configurations respectively. We use a combination of the Metropolis and Wolff algorithms to determine this value.

@@

\subsection{Autocorrelation times and thermalization}
One important aspect is the correlation between different states of the Markov chain. Ideally, each sample of the lattice would be completely independent, but this is not the case. Therefore, we must sweep the lattice sufficiently to produce an effective Monte Carlo result. We use Wolff's automatic windowing procedure \cite{wolff2007} and the magnetic susceptibility $\chi_m$ to estimate the autocorrelation and determine an appropriate number of sweep between measurements for each theory. Furthermore, the initial values of our Markov chain will be highly correlated with the initial random start. Therefore, we also wait a set number of sweeps before taking any measurements so that the lattice can approach a minimum of the action.

@@

\section{Topological charge with a $\theta$ term}

>>>>

\section{Monte Carlo Simulations}
\label{sec:mc}
We implement a Markov Chain Monte Carlo method following Schaich's thesis \cite{schaich2006}. This implementation utilizes a ``random walk,'' i.e. a set of random steps through phase space, to determine statistical values such as correlation functions across the lattice. By the definition of the Markov chain, the probability of adoption of each state, and therefore its inclusion in the Monte Carlo calculation, depends only on the current state and the proposed state. This probability is denoted as $P(\mu\rightarrow\nu)$ where $\mu$ and $\nu$ are the existing and proposed lattice configurations respectively.

@@

\subsection{Thermalization}
Since we begin with a random lattice, the first few configurations of our Monte Carlo simulation will be far from the lattice minimum. We plot the action as a function of metropolis sweeps in Fig.~\ref{fig:therm}.
\begin{figure}[h]
  \centering
      \begin{subfigure}[b]{0.5\textwidth}\centering
        \includegraphics[width=0.6\textwidth]{imgs/therm24.png}
        \caption{$L=24$}
      \end{subfigure}%
      \hfill
      \begin{subfigure}[b]{0.5\textwidth}\centering
        \includegraphics[width=0.6\textwidth]{imgs/therm404.png}
        \caption{$L=404$}
      \end{subfigure}
      \hfill
      \caption{\label{fig:therm} Plots of the action as a function of Monte Carlo time, starting with a random NLSM lattice.}
\end{figure}
Based on this plot, we determine that 1000 sweeps will give sufficient time for the system to reach the classical action minimum. We use this value for the remainder of this study.
\subsection{Autocorrelation times and thermalization}
One important aspect is the correlation between different states of the Markov chain. Ideally, each sample of the lattice would be completely independent, but this is not the case. Since each configuration is based on previous configurations, there is a correlation between each pair of members in the Markov chain, decreasing exponentially based on the number of steps between. Explicitly, the autocorrelation scales as \[e^{t/\tau_{int}}\] where $t$ is the number of steps between configurations and $\tau_{int}$ is the autocorrelation time.\footnote{Though it is called a time, $\tau_{int}$ is in units of Markov Chain steps.} When performing simulations of the lattice, the number of sweeps between measurements should be much larger than $\tau_{int}$.

    We use Wolff's automatic windowing procedure \cite{wolff2007} and the magnetic susceptibility $\chi_m$ to estimate the autocorrelation. Using Wolff's public MatLab code\footnote{\url{https://www.physik.hu-berlin.de/de/com/ALPHAsoft}}, we estimate the autocorrelation time for $L=24$ and $L=404$ lattices. This algorithm identifies the optimal window size with which to calculate the autocorrelation time. We perform this process on $L=24$ and $L=404$ lattices using a thermalization of $1000$ sweeps and $500$ total measurements. Note that this calculation included a Wolff cluster algorithm every $5$ sweeps. The result of this calculation is shown in Fig.~\ref{fig:tauint}.
\begin{figure}[h]
  \centering
      \begin{subfigure}[b]{0.5\textwidth}\centering
        \includegraphics[width=0.6\textwidth]{imgs/tauint24.png}
        \caption{$L=24$, $\tau_{int}=4.43$}
      \end{subfigure}%
      \hfill
      \begin{subfigure}[b]{0.5\textwidth}\centering
        \includegraphics[width=0.6\textwidth]{imgs/tauint404.png}
        \caption{$L=404$, $\tau_{int}=12.81$}
      \end{subfigure}
      \hfill
      \caption{\label{fig:tauint} Plots of automatic windowing procedure used to calculate $\tau_{int}$ for the NLSM model.}
\end{figure}

Based on these two values for $\tau_{int}$, we decide to measure every $50$ sweeps for each simulation. This value will ensure that each measurement is effectively independent.

@@

\section{Topological charge with a $\theta$ term}
In Sec.~\ref{sec:topological charge}, we discussed the introduction of a $\theta$ term into the action. This change made the theory ``topological'', pushing $\langle Q \rangle$ away from zero. In order to calculate $\langle Q \rangle$ as a function of $Q$, we consider the path integral
\begin{align}
    \langle Q \rangle_\theta &=\int \mathcal{D}\e\:Q[\e]e^{-S[\e]+i\theta Q[\e]} \\
                             &=\int \mathcal{D}\e\:\left( Q[\e]e^{i\theta Q[\e]} \right) e^{-S[\e]} \\
                             &=\langle Q e^{i \theta Q} \rangle_{\theta=0}.
\end{align}
Therefore, we can calculate $\langle Q \rangle_\theta$ for arbitrary $\theta$ using the same simulation framework as the $\theta=0$ case.

We can also relate this function to the topological susceptibility. By expanding the exponent as a Taylor series around $\theta=0$, we find that  
\begin{equation}
    \langle Q \rangle_\theta = \langle Q \rangle_{\theta=0} + i \theta \langle Q^2 \rangle_{\theta=0} + O(\theta^2),
\end{equation}
such that 
\begin{align}
    -i \frac{\partial}{\partial \theta} \langle Q \rangle_\theta &= \langle Q^2 \rangle_{\theta=0}\\
                                                              &= \chi_t L^2.
\end{align}
We expect to see the slope of this function at zero increase for large lattice sizes.
